---
title: "examMarks Package"
author: "Andrew Davis, Andriani Hadjiconstanti, and Zamfirescu Ana Maria"
date: "26/4/2017"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

## Package Overview

```{r message = FALSE}
library(examMarks)
```

* This package is designed to mark student answers to based on given exam answer keys and to analysis the distributions of the marks based on different factors.
* It is also capable of generating random student answers and randomly generated answer keys.

The package is split into 3 parts with helpful datasets included as well.

1. Random Generation of Answers
2. Marking student Answers
3. Analysing the Distribution of the Marks

## Datasets

- **students** - dataframe with the ID of students and their respective degree course
```{r message = FALSE}
head(students)
```

- **exams** - dataframe that defines which modules each degree course takes
```{r message = FALSE}
exams
```

## Datasets

- **questions** -  dataframe that lists how many questions a student is asked in each exam and how many total questions each exam has
```{r message = FALSE}
head(questions)
```

- **keys** - list with a sample answer key for each exam
```{r message = FALSE}
summary(keys)
```

## Random Generation of Answers

- To generate an answer sheet for a single student use **generateStudentAnswersForExam()**
- students can answer either a, b, c, d, e, or not at all (N/A) for each question
- if writeToFile == TRUE, then studentID and moduleID must be supplied to create the filename

```{r message = FALSE}
head(generateStudentAnswersForExam(totalNumberofQuestions = 100, 
                                   numberOfQuestionsToAnswer= 30, 
                                   writeToFile = F))
```

## Random Generation of Answers

* To generate all student answer for a given module use **generateAllStudentAnswersForExam()**
* The default values are set to use the provided datasets
    + these can be manipulated or new datasets or files can be used for individual use
    + use readFromFiles to determine whether or not the arguments should be read as Rdata or files
    + use degreeNames if the degree courses are not called 'Biological Sciences' and 'Genetics'
* if writeToFile == FALSE
    + data is output as a list with 2 elements, the students that took the exam and a list of the answers for each student
* if writeToFile == TRUE
    + a file listing the students who took the exam is created
    + a folder is created with the student answer files within it
* if a module is optional for a given degree, course then students are selected at random to take it, with more students taking it being more likely

## Random Generation of Answers

* The student list as the first list element
```{r message = FALSE}
test = generateAllStudentsAnswersForExam('BS281', writeToFile = FALSE)
head(test[[1]])
```
* The next element is the list of dataframes 
    + each student's answers is its own element of the list
* each generates the same output as generateStudentAnswersForExam() for each student's answers

```{r message = FALSE, eval=FALSE}
test = generateAllStudentsAnswersForExam('BS281', writeToFile = FALSE)
test[[1]][2]
```

## Random Generation of Answers

* Can also use **createAnswerKey()** to randomly generate an exam key

```{r message = FALSE}
head(createAnswerKey(numberOfQuestions = 30, writeToFile = FALSE,
                     ansOptions = letters[1:5]))
```

* if writing to a file moduleID is required in order to name the answer key

```{r message = FALSE}
createAnswerKey(numberOfQuestions = 30, writeToFile = TRUE, moduleID = 'BS281',
                ansOptions = letters[1:5])
```

## Marking Students


* To mark all the students for one exam **markStudentsForExam()**
* This function relies on all relevant files to be in the same directory
    + these include the correct answer files, "number_of_questions.tsv" as well as the folders containing the student answer files for each exam
* The output is in the form of a dataframe with studentID, degree course, and mark as columns

```{r message = FALSE}
head(markStudentsForExam(fileDir = './', ExamFilesDir = './BS281studentAnswerFiles',
                    ModuleID = 'BS281'))
```

## Marking Students


* as part of the function that will mark all of the exams degree marks need to be added as well
* the **getDegree()** function is used for that
* In this function a mark is entered and it is converted to a degree

```{r message = FALSE, echo=FALSE}
marks = markStudentsForExam(fileDir = './', ExamFilesDir = './BS281studentAnswerFiles',
                    ModuleID = 'BS281')
```
```{r message = FALSE}
head(addDegrees(marks))
```

## Marking Students

* getDegree() is then used in the **addDegrees()** function to add degrees to each student
* The function takes a dataframe with columns of studentID, degree course, and mark as input
* The same dataframe is outputted with an added column including degree

```{r message = FALSE}
getDegree(65)
```

## Marking Students


* To mark the student answers for all the exams use the **markStudents()** function
* This function takes the directory where everything is located as input
    + This function uses markStudentsForExam(), so all relevant files must be in the same directory
* There is also the option to have the output show on the console or be written into a folder using the writeToFile parameter
* If writeToFile = FALSE, a list of dataframes is created with each one representing the marked answers for a given exam

```{r message = FALSE}
summary(markStudents(fileDir = './'))
```

* If writeToFile = TRUE, a folder is created with a file for each exams marks

## Marking Students

* Using the function **examHist()**, a histogram can be generated using a list containing dataframes for each moduleID  and the moduleID being assessed
* each list dataframe should have columns of studentID, degree course, and mark

```{r message = FALSE, echo=FALSE}
studentID = 1:100
mark = sample(1:100, 100, replace = T)
course = sample(c('Biological Sciences', 'Genetics'), 100, replace = T)
testMarks = data.frame(studentID, course, mark)
testData = list(testMarks)
names(testData) = 'BS281'
```
```{r message = FALSE}
examHist(testData, 'BS281')
```

## Data Analysis

* Using the function **moduleHist()**, a histogram can be generated using a dataframe and the degree subject being assessed
* The dataframe should have columns of studentID, degree course, and mark

```{r message = FALSE}
moduleHist(testMarks, 'Biological Sciences')
```

## Data Analysis

* To analyse the difference in marks between the two degree course within a given module use **testWithinModule()**
* Either a t-test or a wilcoxon test is run
    + A normality test is run to determine which test will be run

```{r message = FALSE}
testWithinModule(testData, 'BS281')
```

## Data Analysis

* To analyse the difference in marks between the two degree course for the overall marks use **testBetweenCourse()**
* Either a t-test or a wilcoxon test is run
    + A normality test is run to determine which test will be run

```{r message = FALSE, echo=FALSE}
studentID = 1:100; mark = sample(1:100, 100, replace = T)
course = sample(c('Biological Sciences', 'Genetics'), 100, replace = T)
testMarks = data.frame(studentID, course, mark)
```
```{r message = FALSE}
testBetweenCourse(testMarks)
```
