---
title: "Team3Package"
author: "Andrew Davis, Andriani Hadjiconstanti, Zamfirescu Ana Maria, and Hannah Leila Tesfay"
date: "26-04-2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r message = FALSE}
library(examMarks)
```

This package is designed to generate random answers to a given exam for a given 
number of students. It can then mark exam answers as well as generating visual 
representaions of the exam marks and analyzing the mark distribution as well. 

## Generating random answers

### generateStudentAnswersForExam()

By using generateStudentAnswersForExam(), one is able to create an answer sheet 
based on the total number of questions in an exam and the amount of questions 
that the student is asked. moduleID and studentID is only required if you are 
writing to a file as they form the file name.

```{r message = FALSE}
head(generateStudentAnswersForExam(totalNumberofQuestions = 100, 
                                     numberOfQuestionsToAnswer= 30, 
                                     writeToFile = F))
```

If one wants to write the generated answer sheet to a file then the file will 
be saved according to moduleID and studentID. For example this file would be 
saved as 'BS281_answers_12.tsv'. 

```{r message = FALSE, eval = FALSE}
generateStudentAnswersForExam(totalNumberofQuestions = 100, 
                              numberOfQuestionsToAnswer= 30, writeToFile = T, 
                              moduleID = 'BS281', studentID = '12')
```

### generateAllStudentsAnswersForExam()

Using the generateAllStudentsAnswersForExam() function one is able to generate 
answer sheets for multiple students. This function uses the 
generateStudentAnswersForExam() function witnin it. Datasets are included in 
this package and are listed as defualt values to demo how the functions runs. 
readFromFiles must equal TRUE if one needs to use external files to run the 
function. 

In this case only moduleID is required. The output is split into a list with the 
first part conatining a list of the students taking the module and the second 
part of the list conatining a list of all the student answers. 

```{r message = FALSE}
test = generateAllStudentsAnswersForExam('BS285', writeToFile = FALSE)
head(test[[1]])
head(data.frame(test[[2]][1]))
```

If one wants to write the answers to files then a folder is created with all of
the student answers contained within. A file is also created that lists the 
students that took the exam.

```{r message = FALSE, eval = FALSE}
generateAllStudentsAnswersForExam('BS285', writeToFile = TRUE)
```

### createAnswerKey()

A randomly generated answer key can also be created using the createAnswerKey() 
function. The number of questions and the answer options must be specified. 

```{r message = FALSE}
head(createAnswerKey(numberOfQuestions = 30, writeToFile = FALSE, 
                     ansOptions = letters[1:5]))
```                           

If the answer key is being written to a file then it is saved according to 
moduleID. For example this file be called 'correct_answers_BS281.dat'

```{r message = FALSE}
createAnswerKey(numberOfQuestions = 30, writeToFile = TRUE, moduleID = 'BS281',
                ansOptions = letters[1:5])
```     

### Datasets provided 

#### exams

A dataframe that indicates what modules each degree course takes, with Yes, No, 
or Optional being listed.

```{r message = FALSE}
head(exams)
``` 

#### keys 

A list containing 5 different exam keys of varying length. All keys have 
possible answers being from "a" to "e".

#### students

A dataframe with the ID and degree course for each student.

```{r message = FALSE}
head(students)
``` 

#### questions

A dataframe with the total number of questions in an exam and the amount of 
questions that need to be answered. 

```{r message = FALSE}
head(questions)
``` 

## Marking Students

### markStudentsForExam()

Using this function a dataframe can be created which contains the marks for each
student for a given exam. The marks are based on the exam answers randomly 
generated from generateAllStudentsAnswersForExam().

```{r message = FALSE}
head(markStudentsForExam(fileDir = './', 
                         ExamFilesDir = './BS281studentAnswerFiles', 
                         ModuleID = 'BS281'))
``` 


### getDegree()

This function is designed to be used in the addDegree() function to add the degree to 
each student. A singular mark is converted to a degree based on the mark.

```{r message = FALSE}
getDegree(61)
```

### addDegrees()

The purpose of this function is to add a degree for each student, amending the 
dataframe as an output.

```{r message = FALSE, echo=FALSE}
studentID = 1:100; mark = sample(1:100, 100, replace = T)
course = sample(c('Biological Sciences', 'Genetics'), 100, replace = T)
testMarks = data.frame(studentID, course, mark)
``` 
```{r message = FALSE}
head(addDegrees(testMarks))
``` 

### markStudents()

Outputs a list of dataframes of student marks for each module or creates a 
folder with files with the marks for each exam depending on if writeToFile is 
TRUE or FALSE.

```{r message = FALSE}
summary(markStudents(fileDir = './'))
``` 

## Data Analysis

### examHist()

A histogram is generated using a list conatining dataframes for each moduleID, with columns of studentID, degree course, and mark, and the moduleID being assesed.

```{r message = FALSE, echo=FALSE}
studentID = 1:100; mark = sample(1:100, 100, replace = T)
course = sample(c('Biological Sciences', 'Genetics'), 100, replace = T)
testMarks = data.frame(studentID, course, mark, sapply(mark, getDegree))
test = list(testMarks); names(test) = 'BS281'
```
```{r message = FALSE}
examHist(test, 'BS281')
``` 

### moduleHist()

A histogram is generated using a dataframe, with columns of studentID, degree 
course, and mark, and the degree subject being assesed.

```{r message = FALSE}
moduleHist(testMarks, 'Biological Sciences')
``` 

### testWithinModule()

Either a t-test or a wilcoxin test is run to test for a difference in the 
distribution of marks between degree courses for a module. A normality test is 
run to determine which test will be run.

```{r message = FALSE}
testWithinModule(test, 'BS281')
``` 

### testBetweenCourse()

Either a t-test or a wilcoxin test is run to test for a difference in the 
distribution of marks between degree courses for overall marks. A normality test 
is run to determine which test will be run.

```{r message = FALSE, echo=FALSE}
studentID = 1:100; mark = sample(1:100, 100, replace = T)
course = sample(c('Biological Sciences', 'Genetics'), 100, replace = T)
testMarks = data.frame(studentID, course, mark, sapply(mark, getDegree))
```
```{r message = FALSE}
testBetweenCourse(testMarks)
``` 

